---
title: "Discovery-Hiring-Analyst-2016"
output: html_document
---
***

# Task

## You must create a reproducible report* answering the following questions:

1. What is our daily overall clickthrough rate? How does it vary between the groups?
2. Which results do people tend to try first? How does it change day-to-day?
3. What is our daily overall zero results rate? How does it vary between the groups?
4. Let session length be approximately the time between the first event and the last event in a session. Choose a variable from the dataset and describe its relationship to session length. Visualize the relationship.
5. Summarize your findings in an executive summary.


# Data

> The dataset comes from a tracking schema that we use for assessing user satisfaction. Desktop users are randomly sampled to be anonymously tracked by this schema which uses a "I'm alive" pinging system that we can use to estimate how long our users stay on the pages they visit. The dataset contains just a little more than a week of EL data.

## Column	Value	Description

1. uuid	string	Universally unique identifier (UUID) for backend event handling.
2. timestamp	integer	The date and time (UTC) of the event, formatted as YYYYMMDDhhmmss.
3. session_id	string	A unique ID identifying individual sessions.
4. group	string	A label ("a" or "b").
5. action	string	Identifies in which the event was created. See below.
6. checkin	integer	How many seconds the page has been open for.
7. page_id	string	A unique identifier for correlating page visits and check-ins.
8. n_results	integer	Number of hits returned to the user. Only shown for searchResultPage events.
9. result_position	integer	The position of the visited page's link on the search engine results page (SERP).

<https://github.com/wikimedia-research/Discovery-Hiring-Analyst-2016>

***

# Import libraries
```{r}

# Import some libraries
library(ggplot2)
library(plyr)
library(dplyr)
library(gridExtra)
library(lubridate)

```


# Import and clean data
```{r}

# Import dataset CSV and explicitly cast each column type

events <- read.table(
  file = "C:/Users/Eirika/Documents/Code/Datasets/events_log/events_log.csv",
  sep = ",", header = TRUE,
  colClasses =  ( 
    c(
      "character", 
      "character", 
      "character", 
      "character", 
      "character",
      "integer", 
      "character",
      "integer", 
      "integer"
    )
  ))

# Using ymd_hms and date, convert timestamp column
events$timestamp <- ymd_hms(events$timestamp)
events$date <- date(events$timestamp)

```

## Troubleshooting

> Want to select all NA events that werent timestamped correctly and figure out what to do with them

```{r}

# Troubleshooting

# Pull all records out with NA timestamp to a separate data frame to view

fails <- events[is.na(events$timestamp),]

# Remove fails from events df?

```

*** 

## 1. What is our daily overall clickthrough rate? How does it vary between the groups?

> The daily clickthrough rate varies from 6654 clickthroughs on 3/1/16 to 3323 clickthroughs on 3/5/16

> Group A makes up the bulk of the clickthrough events, whereas group B has far fewer.

```{r}

# Pull out data only of visitPage clickthrough events

clickthrough <- events[events$action == 'visitPage',]

# Pull out clickthroughs that resulted from non-NA result position
clickthrough <- clickthrough[!is.na(clickthrough$result_position),]

# Chart by day

cpd <- ggplot(clickthrough, aes(x=clickthrough$date)) + geom_bar() + labs(title= "Overall",x="Date", y="Clickthroughs")  + theme(axis.text.x=element_text(angle=90), axis.text.y=element_text(angle=0))
print(cpd)

# Chart per group a and b

groupa <- clickthrough[clickthrough$group == 'a',]
groupb <- clickthrough[clickthrough$group == 'b',]

cpda <- ggplot(groupa, aes(x=groupa$date)) + labs(title= "Group A",x="Date", y="Clickthroughs") + geom_bar() + theme(axis.text.x=element_text(angle=90), axis.text.y=element_text(angle=0))

cpdb <- ggplot(groupb, aes(x=groupb$date)) + labs(title= "Group B",x="Date", y="Clickthroughs") + geom_bar() + theme(axis.text.x=element_text(angle=90), axis.text.y=element_text(angle=0))


grid.arrange(cpd, cpda, cpdb, ncol=3)

# Overall daily clickthrough rate
#count(clickthrough$date)

# Group A clickthrough rate
#count(groupa$date)

# Group B clickthrough rate
#count(groupb$date)

# How large is each group?
#groupsize <- count(events$group)
#print(groupsize)
#Error in UseMethod("groups") : no applicable method for 'groups' applied to an object of class "character"
```

***

## 2. Which results do people tend to try first? How does it change day-to-day?

> By far the most people try the first result first. On 3/1/16, there are a cluster of search results that were around position 4000. This was not the case on any other day. 3/4/16 had the largest variability in search result position.

```{r}

# Pull out all data that have a results_position that is not null

results <- events[!(is.na(events$result_position)),]

# count(results$result_position) # for freq table

# Graph of distribution of results_position
rtf <- ggplot(results, aes(x=results$result_position)) + geom_bar() + coord_cartesian(xlim=c(0,10)) + labs(title = "Results Clicked First", x="Result Position", y= "Frequency")

print(rtf)

# Zoom in on only result positions in the top 500

results_500 <- results[results$result_position < 500,]

days_500 <- ggplot(results_500, aes(x=results_500$date, y=results_500$result_position, colour = results_500$result_position)) + geom_jitter() + labs(title = "Results Clicked First", x="Day", y= "Result Position", colour = "Result Position") + scale_colour_gradientn(colours=rainbow(4))

print(days_500)

# Zoom in on only result positions in the top 200

results_200 <- results_500[results_500$result_position < 200,]

days_200 <- ggplot(results_200, aes(results_200$date, y=results_200$result_position, colour = results_200$result_position)) + geom_jitter() + labs(title = "Results Clicked First", x="Day", y= "Result Position", colour = "Result Position") + scale_colour_gradientn(colours=rainbow(4))

print(days_200)

```

***

## 3. What is our daily overall zero results rate? How does it vary between the groups?

> The daily overall zero results rate ranges between 10 and 15% each day. Group A is closer to the overall trend whereas Group b as a population displays less variability in zero results rate.

```{r}

# Pull out only search result page action
search_results <- events[events$action == "searchResultPage",]

# Pull out only n_results of 0, exclude N/A

not_na_results <- search_results[!is.na(search_results$n_results),]

zero_results <- not_na_results[not_na_results$n_results == 0,]

# Pull out group b and group A data

zero_a <- zero_results[zero_results$group == "a",]

zero_b <- zero_results[zero_results$group == "b",]

# Plot by day

zrpd <- ggplot(zero_results) + geom_bar(mapping = aes(x=zero_results$date,y =..prop..), stat = "count", group = 1) + labs(title = "Frequency of Zero Results Per Day", x="Day", y= "Percentage of zero Results") + scale_y_continuous(labels = scales::percent_format())

# Plot by group

zra <- ggplot(zero_a) + geom_bar(mapping = aes(x=zero_a$date,y =..prop..), stat = "count", group = 1) + labs(title = "Frequency of Zero Results Per Day", x="Day", y= "Percentage of zero Results") + scale_y_continuous(labels = scales::percent_format())

zrb <- ggplot(zero_b) + geom_bar(mapping = aes(x=zero_b$date,y =..prop..), stat = "count", group = 1) + labs(title = "Frequency of Zero Results Per Day", x="Day", y= "Percentage of zero Results") + scale_y_continuous(labels = scales::percent_format())

print(zrpd)
print(zra)
print(zrb)

```

## 4. Let session length be approximately the time between the first event and the last event in a session. Choose a variable from the dataset and describe its relationship to session length. Visualize the relationship.

> Session length between group a and b appears to be slightly different, with group b having a smaller median session length.


```{r}

# Group by session id and generate a new column: len that is the difference between the earliest and latest timestamps in a session (in s)
# Remove lengths of 0

sessions <- events %>%
  select(group,session_id,timestamp,date) %>%
  group_by(group,session_id) %>%
  summarise(min=min(timestamp),max=max(timestamp)) %>%
  mutate(len = max - min) %>%
  filter(len!=0)

# Box plot comparing groups a and b session length variable
# Remove lengths greater than the 95th percentile

box <- ggplot(sessions %>% filter(len < quantile(sessions$len, 0.95)), aes(factor(group), len)) + geom_boxplot(aes(fill = factor(group)), width=0.5, outlier.colour = "dodgerblue", outlier.size = 1, outlier.shape = 16, outlier.stroke = 1) + labs(title="Session length in group a vs group b", x="Group", y="Session Length (s)")

print(box)

# Box plot comparing days to session length, separated by group

```

